Implementing Equational Constraints\\ in a Functional Language

### 2011-07-01 papers/wlp2011/paper.tex

\begin{tabular}{|l|c|c|c|}
\hline
Expression: &\code{==} & \code{=:=} & \code{=:<=}\\
\hline
\code{last (take 100000 (repeat failed) ++ [1])} & \code{failed} & \code{failed} & 0.14\\ 
\code{last (map (inc 0) [1..100000])} & stack overflow & 0.45 & 0.13\\
\code{simplify}                       &10.34&10.33&7.27\\
\code{varsInExp} & 0.49& 84.92 & 0.09\\
\code{toPeano 5000 == halfS (toPeano 10000)} & 23.3 & 1.81 & no Fun Pat\\
\hline
\end{tabular}

The table above shows the evaluation of our approach. Primarily we
used the same Benchmarks as in \cite{}

Benchmarks for Last with ==, =:= and func. patterns

also short comparison of =:= with PAKCS and MCC

>>>>

Unlike the Curry implementations PAKCS and MCC our approach does
not manipulate a global state to unify a logic variable. Nevertheless,
our benchmarks show that it can in most cases
compete or even outperform these implementations.
The programs used for the benchmarks are primarily the same as in \cite{AntoyHanus05LOPSTR}.

\begin{tabular}{l l}
\bf{last}& Computes the last element of a list using strict equality\\
\bf{simplify}& Simplification of symbolic arithmetic expressions\\
\bf{varInExp}& Nondeterministically returns a variable occuring
                in a symbolic arithmetic expression \\
\bf{half}& Computes the half of a peano number using logic variables\\
\end{tabular}

The table below shows a comparison of the curry systems where
strict equality (\code{=:=}) is used for unification:
\footnote{\ccode{inc x n} is a naive addition that adds 1 \code{n} times
          to its argument x }

\begin{tabular}{|l|c|c|c|}
\hline
Expression: & KiCS2 & PAKCS & MCC \\
\code{last (map (inc 0) [1..100000])} &0.52 & 4.27&0.15 \\
\code{grep} & 0.13& 2.83& 0.15\\
\code{simplify} &10.33&0.15&0.01\\
\code{varsInExp} & 86.27 & 0.35  & 0.03\\
\code{fromPeano (half (toPeano 10000))} & &  & \\
\code{palindrom} & 13.45 &32.82 & 1.04\\
\code{horseman} & & &\\
\hline
\end{tabular}

Some of our examples can also be defined with function patterns, i.e.,
the lazy equality operator \code{=:<=} is used. The sole other
Curry system featuring function patterns is PAKCS:

\begin{tabular}{|l|c|c|c|}
\hline
Expression: & KiCS2 & PAKCS \\
\code{last (map (inc 0) [1..100000])} & 0.20& 3.93\\
\code{simplify}  &7.09&0.26&\\
\code{varsInExp} & 0.14 & 0.69\\
\code{fromPeano (half (toPeano 10000))} & &  \\
\code{palindrom} & 17.55 & $\infty$\\
\code{horseman} & & \\
\hline
\end{tabular}

In Section \ref{sec:Unification} we stated that equational
constraits can be implemented using ordinary pattern matching.
We also mentioned that such an implementation would lead to
a large search space that has to be traversed to extract
the results of a computation. To show the benificial effects
of our equality constraint implementations we compare them to
the results using the equality operator \code{==} that
produces such a search space:

\begin{tabular}{|l|c|c|c|}
\hline
Expression: &\code{==} & \code{=:=} & \code{=:<=}\\
\hline 
\code{last (map (inc 0) [1..100000])} & 118.33 & 0.52 & 1.40 \\
\code{grep} & 0.94 & 0.13 &\\
\code{simplify} & 9.18& 10.33 & 7.09\\
\code{varsInExp} & 0.50& 86.27 & 0.14\\
\code{fromPeano (half (toPeano 10000))} &  &  & \\
\code{palindrom} & 31.12&13.45 & 17.55\\
\code{horseman} & & & \\
\hline
\end{tabular}
